% $Id$

\begin{abstract}
The Component Workload Emulator (CoWorkEr) Utilization Test Suite (also known  
as CUTS) is a model-driven engineering (MDE)-based tool that allows developers 
to conduct system integration test at early stages of development. Furthermore,
is used to help identify system integration problems, such as failure to meet 
end-to-end response time due to poor infrastructure implementation, before they 
become to hard to locate and resolve later in the development lifecycle. 

Developers and testers use CUTS to model the expected behavior of their real 
components and the structure of their system. The behavior models are then 
used to generate emulation components that are compliant with the target 
technologies specification and have the same interfaces and attributes as their 
real counterpart. The emulated components are then executed in the target, or 
representative, environment and different QoS metrics (\textit{e.g.}, throughput, 
response time, and service rate) are monitored and collected. 

Developers and testers can grapically view the collected metrics while the system 
is executing to understand its current performance, and gain insight on how to 
improve it. Developer also have the option of viewing and analyzing collected 
performance metrics postmortem. Lastly, as the real component's implementation 
is completed, it can replace its respective emulation component to produce more 
realistic results and facilitate continuous system integration.

This book is the user's guide to CUTS. It contains information about building
and installing CUTS, using its modeling languages to create system experiments,
and details on CUTS many tools that help simpify component-based distributed
system development, not only during the early stages of development, but 
throughout the entire development lifecycle.
\end{abstract}
